---
title: "Model description"
author: "Michael Bedward, Karen Ross"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Description of the simulation}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

The woodland package contains a demographic stand model in which growth, survival and recruitment are influenced by rainfall, competition and disturbance (fire, grazing or tree removal). Different woodland species or functional types can be modelled as pure or mixed stands. The model follows cohorts of trees in a single stand occupying an environmentally homogeneous one hectare site. 

Each cohort is described by its species, age, height and the number of individuals. Other attributes (canopy radius and crown spheroid surface area), are calculated from tree height using deterministic relationships provided to the model as a look-up table. Canopy radius is used to calculate core canopy area (the central 2-D projected area of a canopy) which influences recruitment (and fire) in the model. Core canopy area is derived using a parameter overlap which defines the outer portion of the canopy radius that may be overlapped by the canopies of new recruits. Resource use is the contribution of each individual tree of a given height to total resource use or competition in the stand which influences growth and survival in the model. We use crown spheroid surface area (surface areas of the oblate spheroids depicting each crown summed over all cohorts), as a surrogate for total resource demand (competition) in the stand. Rainfall and competition constraints on growth and survival are applied using inverse logit functions of the form:

$$
x_t = logit^{-1}( \alpha + \beta_0 * radius_t + \beta_1 * {height_t}^{\beta_2} )^{\beta_3}
$$

## Fire

### Fire patchiness

By default, each fire in the simulation burns all trees in all cohorts, regardless of the realized fire intensity. In reality, low intensity fires are more likely to be patchy, leaving some trees unburnt. Fire patchiness can be included in the simulation by providing a user-defined function which maps values of realized fire intensity to the proportion of trees to burn. The function is passed to the simulation via the `fire.patchiness.func` argument. 

As an example, this function maps intensity values to proportion burnt using a Burr type XII cumulative distribution function (`pburr` and related functions for the Burr distribution are included in the woodland package):

```{r}
patchiness <- function( fire.intensity ) {
  # using a Burr XII distribution to map non-zero intensity to 
  # proporition burnt
  p <- woodland::pburr(fire.intensity, c=2, k=1, scale=1.5)
  
  # zero intensity values map to zero proportion
  p[ fire.intensity == 0 ] <- 0
  
  p
}
```

```{r, echo=FALSE, fig.width=5}
library(ggplot2)

x <- seq(0, 10, 0.1)
dat <- data.frame(intensity = x, prop = patchiness(x))

ggplot(dat, aes(intensity, prop)) + 
  geom_line() +
  theme_bw() +
  labs(x="fire intensity", y="proportion burnt")


```


Once the proportion of the stand which will be burnt $p_{burn}$ has been derived from the fire intensity, the overall probability of an individual tree surviving a fire $p_{survival}$ is given by the probability that it escapes fire $(1 - p_{burn})$ plus the probability that it is burnt but survives $p_{burn} \times p_{fire.survival}$.

$$ 
\begin{aligned}
p_{survival} & = 1 - p_{burn} + p_{burn}p_{fire.survival} \\
             & = 1 - p_{burn}(1 + p_{fire.survival})
\end{aligned}
$$

For a cohort of $N_{initial}$ trees, a random number of trees surviving a fire, either by escaping fire or being burnt but surviving, can then be drawn from a Binomial distribution:

$$
N_{surviving} \sim Binomial(N_{initial}, p_{survival}) 
$$

